---
title: "Clara"
author: "Anqi Chi SID:460204008"
date: "2018年10月18日"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(tidyverse)
```


```{r echo=FALSE}
#Read data
dat_orig <- read.csv("nutmstatData2018.csv",header=TRUE)
dat <-  dat_orig
colNames <- colnames(dat_orig)

#Recode variables
# for (i in 1:length(colNames)) { cat("categoricalList[",i,"] <- FALSE # ",colNames[i],"\n"); }

categoricalList <- c()
categoricalList[ 1 ] <- FALSE #  BMISC 
categoricalList[ 2 ] <- FALSE #  AGEC 
categoricalList[ 3 ] <- TRUE #  SMSBC 
categoricalList[ 4 ] <- TRUE #  COBBC 
categoricalList[ 5 ] <- TRUE #  FEMLSBC 
categoricalList[ 6 ] <- FALSE #  PHDKGWBC 
categoricalList[ 7 ] <- FALSE #  PHDCMHBC 
categoricalList[ 8 ] <- FALSE #  EXLWTBC 
categoricalList[ 9 ] <- FALSE #  EXLWMBC 
categoricalList[ 10 ] <- FALSE #  EXLWVBC 
categoricalList[ 11 ] <- FALSE #  PHDCMWBC 
categoricalList[ 12 ] <- FALSE #  BMR 
categoricalList[ 13 ] <- FALSE #  EIBMR1 
categoricalList[ 14 ] <- TRUE #  SF2SA1QN 
categoricalList[ 15 ] <- TRUE #  INCDEC 
categoricalList[ 16 ] <- TRUE #  DIABBC 
categoricalList[ 17 ] <- TRUE #  HCHOLBC 
categoricalList[ 18 ] <- TRUE #  HSUGBC 
categoricalList[ 19 ] <- TRUE #  HYPBC 
categoricalList[ 20 ] <- FALSE #  ENERGYT1 
categoricalList[ 21 ] <- FALSE #  ENRGYT1 
categoricalList[ 22 ] <- FALSE #  MOISTT1 
categoricalList[ 23 ] <- FALSE #  PROTT1 
categoricalList[ 24 ] <- FALSE #  FATT1 
categoricalList[ 25 ] <- FALSE #  CHOWSAT1 
categoricalList[ 26 ] <- FALSE #  CHOWOAT1 
categoricalList[ 27 ] <- FALSE #  STARCHT1 
categoricalList[ 28 ] <- FALSE #  SUGART1 
categoricalList[ 29 ] <- FALSE #  FIBRET1 
categoricalList[ 30 ] <- FALSE #  ALCT1 
categoricalList[ 31 ] <- FALSE #  PREVAT1 
categoricalList[ 32 ] <- FALSE #  PROVAT1 
categoricalList[ 33 ] <- FALSE #  RETEQT1 
categoricalList[ 34 ] <- FALSE #  B1T1 
categoricalList[ 35 ] <- FALSE #  B2T1 
categoricalList[ 36 ] <- FALSE #  B3T1 
categoricalList[ 37 ] <- FALSE #  NIACINT1 
categoricalList[ 38 ] <- FALSE #  FOLATT1 
categoricalList[ 39 ] <- FALSE #  FOLICT1 
categoricalList[ 40 ] <- FALSE #  TOTFOLT1 
categoricalList[ 41 ] <- FALSE #  FOLEQT1 
categoricalList[ 42 ] <- FALSE #  B6T1 
categoricalList[ 43 ] <- FALSE #  B12T1 
categoricalList[ 44 ] <- FALSE #  VITCT1 
categoricalList[ 45 ] <- FALSE #  VITET1 
categoricalList[ 46 ] <- FALSE #  CALCT1 
categoricalList[ 47 ] <- FALSE #  IODINET1 
categoricalList[ 48 ] <- FALSE #  IRONT1 
categoricalList[ 49 ] <- FALSE #  MAGT1 
categoricalList[ 50 ] <- FALSE #  PHOST1 
categoricalList[ 51 ] <- FALSE #  POTAST1 
categoricalList[ 52 ] <- FALSE #  SELT1 
categoricalList[ 53 ] <- FALSE #  SODIUMT1 
categoricalList[ 54 ] <- FALSE #  ZINCT1 
categoricalList[ 55 ] <- FALSE #  CAFFT1 
categoricalList[ 56 ] <- FALSE #  CHOLT1 
categoricalList[ 57 ] <- FALSE #  SATFATT1 
categoricalList[ 58 ] <- FALSE #  MUFATT1 
categoricalList[ 59 ] <- FALSE #  PUFATT1 
categoricalList[ 60 ] <- FALSE #  LAT1 
categoricalList[ 61 ] <- FALSE #  ALAT1 
categoricalList[ 62 ] <- FALSE #  LCN3T1 
categoricalList[ 63 ] <- FALSE #  TRANST1 
categoricalList[ 64 ] <- FALSE #  PROPER1 
categoricalList[ 65 ] <- FALSE #  FATPER1 
categoricalList[ 66 ] <- FALSE #  LAPER1 
categoricalList[ 67 ] <- FALSE #  ALAPER1 
categoricalList[ 68 ] <- FALSE #  CHOPER1 
categoricalList[ 69 ] <- FALSE #  SUGPER1 
categoricalList[ 70 ] <- FALSE #  STARPER1 
categoricalList[ 71 ] <- FALSE #  ALCPER1 
categoricalList[ 72 ] <- FALSE #  SATPER1 
categoricalList[ 73 ] <- FALSE #  TRANPER1 
categoricalList[ 74 ] <- FALSE #  FIBRPER1 
categoricalList[ 75 ] <- FALSE #  MONOPER1 
categoricalList[ 76 ] <- FALSE #  POLYPER1 
categoricalList[ 77 ] <- FALSE #  ADTOTSE 
categoricalList[ 78 ] <- TRUE #  BDYMSQ04 
categoricalList[ 79 ] <- FALSE #  DIASTOL 
categoricalList[ 80 ] <- TRUE #  DIETQ12 
categoricalList[ 81 ] <- TRUE #  DIETQ14 
categoricalList[ 82 ] <- TRUE #  DIETQ5 
categoricalList[ 83 ] <- TRUE #  DIETQ8 
categoricalList[ 84 ] <- TRUE #  DIETRDI 
categoricalList[ 85 ] <- TRUE #  SABDYMS 
categoricalList[ 86 ] <- TRUE #  SEX 
categoricalList[ 87 ] <- FALSE #  SLPTIME 
categoricalList[ 88 ] <- TRUE #  SMKDAILY 
categoricalList[ 89 ] <- TRUE #  SMKSTAT 
categoricalList[ 90 ] <- FALSE #  SYSTOL 
categoricalList[ 91 ] <- TRUE #  ABSPID 
categoricalList[ 92 ] <- FALSE #  GRAINS1N 
categoricalList[ 93 ] <- FALSE #  WHOLGR1N 
categoricalList[ 94 ] <- FALSE #  REFGRA1N 
categoricalList[ 95 ] <- FALSE #  VEGLEG1N 
categoricalList[ 96 ] <- FALSE #  GREENS1N 
categoricalList[ 97 ] <- FALSE #  VGORSV1N 
categoricalList[ 98 ] <- FALSE #  STARCH1N 
categoricalList[ 99 ] <- FALSE #  LEGVEG1N 
categoricalList[ 100 ] <- FALSE #  OTHVEG1N 
categoricalList[ 101 ] <- FALSE #  FRUIT1N 
categoricalList[ 102 ] <- FALSE #  FRJUIC1N 
categoricalList[ 103 ] <- FALSE #  DAIRY1N 
categoricalList[ 104 ] <- FALSE #  MEAT1N 
categoricalList[ 105 ] <- FALSE #  RDMTL1N 
categoricalList[ 106 ] <- FALSE #  RDMTLU1N 
categoricalList[ 107 ] <- FALSE #  RDMTN1N 
categoricalList[ 108 ] <- FALSE #  RDMTNU1N 
categoricalList[ 109 ] <- FALSE #  RDMTNP1N 
categoricalList[ 110 ] <- FALSE #  PLTYL1N 
categoricalList[ 111 ] <- FALSE #  PLTYLU1N 
categoricalList[ 112 ] <- FALSE #  FISH1N 
categoricalList[ 113 ] <- FALSE #  EGGS1N 
categoricalList[ 114 ] <- FALSE #  LEGMT1N 
categoricalList[ 115 ] <- FALSE #  NUTS1N 
categoricalList[ 116 ] <- FALSE #  WATER1N 
categoricalList[ 117 ] <- FALSE #  UNSAT1N 
categoricalList[ 118 ] <- FALSE #  FRESUG1N 
categoricalList[ 119 ] <- FALSE #  ADDSUG1N 
categoricalList[ 120 ] <- FALSE #  WATERG1N 
categoricalList[ 121 ] <- FALSE #  PEFRESD1 
categoricalList[ 122 ] <- FALSE #  PEADDSD1 
categoricalList[ 123 ] <- TRUE #  ALTNTR 
categoricalList[ 124 ] <- TRUE #  ALTRESB 
categoricalList[ 125 ] <- TRUE #  APOBNTR 
categoricalList[ 126 ] <- TRUE #  APOBRESB 
categoricalList[ 127 ] <- TRUE #  B12RESB 
categoricalList[ 128 ] <- TRUE #  BIORESPC 
categoricalList[ 129 ] <- TRUE #  CHOLNTR 
categoricalList[ 130 ] <- TRUE #  CHOLRESB 
categoricalList[ 131 ] <- TRUE #  CVDMEDST 
categoricalList[ 132 ] <- TRUE #  DIAHBRSK 
categoricalList[ 133 ] <- TRUE #  FASTSTAD 
categoricalList[ 134 ] <- TRUE #  FOLATREB 
categoricalList[ 135 ] <- TRUE #  GGTNTR 
categoricalList[ 136 ] <- TRUE #  GGTRESB 
categoricalList[ 137 ] <- TRUE #  GLUCFPD 
categoricalList[ 138 ] <- TRUE #  GLUCFREB 
categoricalList[ 139 ] <- TRUE #  HBA1PREB 
categoricalList[ 140 ] <- TRUE #  HDLCHREB 
categoricalList[ 141 ] <- TRUE #  LDLNTR 
categoricalList[ 142 ] <- TRUE #  LDLRESB 
categoricalList[ 143 ] <- TRUE #  TRIGNTR 
categoricalList[ 144 ] <- TRUE #  TRIGRESB 

for (i in 1:length(categoricalList)) {
  if (categoricalList[ i ]) {
      dat[,i] <- as.factor(dat[  ,i])
  }
}

# for (i in 1:length(colNames)) { cat("excludeList[[",i,"]] <- c() # ",colNames[i],"\n"); }

excludeList <- list()

excludeList[[ 1 ]] <- c(0,98,99) #  BMISC 
excludeList[[ 2 ]] <- c() #  AGEC 
excludeList[[ 3 ]] <- c(0) #  SMSBC 
excludeList[[ 4 ]] <- c() #  COBBC 
excludeList[[ 5 ]] <- c() #  FEMLSBC 
excludeList[[ 6 ]] <- c(0,997,998,999) #  PHDKGWBC 
excludeList[[ 7 ]] <- c(0,998,999) #  PHDCMHBC 
excludeList[[ 8 ]] <- c(9996,9999) #  EXLWTBC 
excludeList[[ 9 ]] <- c(9996, 9999) #  EXLWMBC 
excludeList[[ 10 ]] <- c(9996 , 9999) #  EXLWVBC 
excludeList[[ 11 ]] <- c(0,998,999) #  PHDCMWBC 
excludeList[[ 12 ]] <- c(99998) #  BMR 
excludeList[[ 13 ]] <- c(998) #  EIBMR1 
excludeList[[ 14 ]] <- c(0,99) #  SF2SA1QN 
excludeList[[ 15 ]] <- c(0,98,99) #  INCDEC 
excludeList[[ 16 ]] <- c() #  DIABBC 
excludeList[[ 17 ]] <- c() #  HCHOLBC 
excludeList[[ 18 ]] <- c() #  HSUGBC 
excludeList[[ 19 ]] <- c() #  HYPBC 
excludeList[[ 20 ]] <- c() #  ENERGYT1 
excludeList[[ 21 ]] <- c() #  ENRGYT1 
excludeList[[ 22 ]] <- c() #  MOISTT1 
excludeList[[ 23 ]] <- c() #  PROTT1 
excludeList[[ 24 ]] <- c() #  FATT1 
excludeList[[ 25 ]] <- c() #  CHOWSAT1 
excludeList[[ 26 ]] <- c() #  CHOWOAT1 
excludeList[[ 27 ]] <- c() #  STARCHT1 
excludeList[[ 28 ]] <- c() #  SUGART1 
excludeList[[ 29 ]] <- c() #  FIBRET1 
excludeList[[ 30 ]] <- c() #  ALCT1 
excludeList[[ 31 ]] <- c() #  PREVAT1 
excludeList[[ 32 ]] <- c() #  PROVAT1 
excludeList[[ 33 ]] <- c() #  RETEQT1 
excludeList[[ 34 ]] <- c() #  B1T1 
excludeList[[ 35 ]] <- c() #  B2T1 
excludeList[[ 36 ]] <- c() #  B3T1 
excludeList[[ 37 ]] <- c() #  NIACINT1 
excludeList[[ 38 ]] <- c() #  FOLATT1 
excludeList[[ 39 ]] <- c() #  FOLICT1 
excludeList[[ 40 ]] <- c() #  TOTFOLT1 
excludeList[[ 41 ]] <- c() #  FOLEQT1 
excludeList[[ 42 ]] <- c() #  B6T1 
excludeList[[ 43 ]] <- c() #  B12T1 
excludeList[[ 44 ]] <- c() #  VITCT1 
excludeList[[ 45 ]] <- c() #  VITET1 
excludeList[[ 46 ]] <- c() #  CALCT1 
excludeList[[ 47 ]] <- c() #  IODINET1 
excludeList[[ 48 ]] <- c() #  IRONT1 
excludeList[[ 49 ]] <- c() #  MAGT1 
excludeList[[ 50 ]] <- c() #  PHOST1 
excludeList[[ 51 ]] <- c() #  POTAST1 
excludeList[[ 52 ]] <- c() #  SELT1 
excludeList[[ 53 ]] <- c() #  SODIUMT1 
excludeList[[ 54 ]] <- c() #  ZINCT1 
excludeList[[ 55 ]] <- c() #  CAFFT1 
excludeList[[ 56 ]] <- c() #  CHOLT1 
excludeList[[ 57 ]] <- c() #  SATFATT1 
excludeList[[ 58 ]] <- c() #  MUFATT1 
excludeList[[ 59 ]] <- c() #  PUFATT1 
excludeList[[ 60 ]] <- c() #  LAT1 
excludeList[[ 61 ]] <- c() #  ALAT1 
excludeList[[ 62 ]] <- c() #  LCN3T1 
excludeList[[ 63 ]] <- c() #  TRANST1 
excludeList[[ 64 ]] <- c() #  PROPER1 
excludeList[[ 65 ]] <- c() #  FATPER1 
excludeList[[ 66 ]] <- c() #  LAPER1 
excludeList[[ 67 ]] <- c() #  ALAPER1 
excludeList[[ 68 ]] <- c() #  CHOPER1 
excludeList[[ 69 ]] <- c() #  SUGPER1 
excludeList[[ 70 ]] <- c() #  STARPER1 
excludeList[[ 71 ]] <- c() #  ALCPER1 
excludeList[[ 72 ]] <- c() #  SATPER1 
excludeList[[ 73 ]] <- c() #  TRANPER1 
excludeList[[ 74 ]] <- c() #  FIBRPER1 
excludeList[[ 75 ]] <- c() #  MONOPER1 
excludeList[[ 76 ]] <- c() #  POLYPER1 
excludeList[[ 77 ]] <- c(9996,9999) #  ADTOTSE 
excludeList[[ 78 ]] <- c(0,6) #  BDYMSQ04 
excludeList[[ 79 ]] <- c(0,998,999) #  DIASTOL 
excludeList[[ 80 ]] <- c(0,6) #  DIETQ12 
excludeList[[ 81 ]] <- c(0,6) #  DIETQ14 
excludeList[[ 82 ]] <- c(0) #  DIETQ5 
excludeList[[ 83 ]] <- c(0) #  DIETQ8 
excludeList[[ 84 ]] <- c(0,3) #  DIETRDI 
excludeList[[ 85 ]] <- c(0,8,9) #  SABDYMS 
excludeList[[ 86 ]] <- c() #  SEX 
excludeList[[ 87 ]] <- c(9998,9999) #  SLPTIME 
excludeList[[ 88 ]] <- c() #  SMKDAILY 
excludeList[[ 89 ]] <- c(0) #  SMKSTAT 
excludeList[[ 90 ]] <- c(0,998,999) #  SYSTOL 
excludeList[[ 91 ]] <- c() #  ABSPID 
excludeList[[ 92 ]] <- c() #  GRAINS1N 
excludeList[[ 93 ]] <- c() #  WHOLGR1N 
excludeList[[ 94 ]] <- c() #  REFGRA1N 
excludeList[[ 95 ]] <- c() #  VEGLEG1N 
excludeList[[ 96 ]] <- c() #  GREENS1N 
excludeList[[ 97 ]] <- c() #  VGORSV1N 
excludeList[[ 98 ]] <- c() #  STARCH1N 
excludeList[[ 99 ]] <- c() #  LEGVEG1N 
excludeList[[ 100 ]] <- c() #  OTHVEG1N 
excludeList[[ 101 ]] <- c() #  FRUIT1N 
excludeList[[ 102 ]] <- c() #  FRJUIC1N 
excludeList[[ 103 ]] <- c() #  DAIRY1N 
excludeList[[ 104 ]] <- c() #  MEAT1N 
excludeList[[ 105 ]] <- c() #  RDMTL1N 
excludeList[[ 106 ]] <- c() #  RDMTLU1N 
excludeList[[ 107 ]] <- c() #  RDMTN1N 
excludeList[[ 108 ]] <- c() #  RDMTNU1N 
excludeList[[ 109 ]] <- c() #  RDMTNP1N 
excludeList[[ 110 ]] <- c() #  PLTYL1N 
excludeList[[ 111 ]] <- c() #  PLTYLU1N 
excludeList[[ 112 ]] <- c() #  FISH1N 
excludeList[[ 113 ]] <- c() #  EGGS1N 
excludeList[[ 114 ]] <- c() #  LEGMT1N 
excludeList[[ 115 ]] <- c() #  NUTS1N 
excludeList[[ 116 ]] <- c() #  WATER1N 
excludeList[[ 117 ]] <- c() #  UNSAT1N 
excludeList[[ 118 ]] <- c() #  FRESUG1N 
excludeList[[ 119 ]] <- c() #  ADDSUG1N 
excludeList[[ 120 ]] <- c() #  WATERG1N 
excludeList[[ 121 ]] <- c() #  PEFRESD1 
excludeList[[ 122 ]] <- c() #  PEADDSD1 
excludeList[[ 123 ]] <- c(0,8) #  ALTNTR 
excludeList[[ 124 ]] <- c(97,98) #  ALTRESB 
excludeList[[ 125 ]] <- c(0,8) #  APOBNTR 
excludeList[[ 126 ]] <- c(97,98) #  APOBRESB 
excludeList[[ 127 ]] <- c(97,98) #  B12RESB 
excludeList[[ 128 ]] <- c(0) #  BIORESPC 
excludeList[[ 129 ]] <- c(0,8) #  CHOLNTR 
excludeList[[ 130 ]] <- c(97,98) #  CHOLRESB 
excludeList[[ 131 ]] <- c(0,8) #  CVDMEDST 
excludeList[[ 132 ]] <- c(0,8) #  DIAHBRSK 
excludeList[[ 133 ]] <- c(0) #  FASTSTAD 
excludeList[[ 134 ]] <- c(97,98) #  FOLATREB 
excludeList[[ 135 ]] <- c(8) #  GGTNTR 
excludeList[[ 136 ]] <- c(97,98) #  GGTRESB 
excludeList[[ 137 ]] <- c(0,8) #  GLUCFPD 
excludeList[[ 138 ]] <- c(97,98) #  GLUCFREB 
excludeList[[ 139 ]] <- c(7,8) #  HBA1PREB 
excludeList[[ 140 ]] <- c(7,8) #  HDLCHREB 
excludeList[[ 141 ]] <- c(0,8) #  LDLNTR 
excludeList[[ 142 ]] <- c(97,98) #  LDLRESB 
excludeList[[ 143 ]] <- c(0,8) #  TRIGNTR 
excludeList[[ 144 ]] <- c(97,98) #  TRIGRESB 

for (i in 1:length(excludeList)) {
  dat[ dat[,i]%in%excludeList[[ i ]],i] <- NA
}

``` 
```{R echo=FALSE}
#Remove any variables with more than 50% NAs
keepR = which(apply(is.na(dat), 1, mean)<= 1) #keep rows where there are less than 100% NAs
keepC = which(apply(is.na(dat), 2, mean) <= 0.5) #keep columns where there are less than 50% NAs
proj_dat = dat[keepR, keepC]
```

#Data Clean
```{r echo=FALSE}
DataC<-proj_dat[proj_dat$AGEC>=18 & proj_dat$FATPER1<=25 & proj_dat$CHOPER1>=55,]

high.carb<-DataC$CHOPER1>55
high.protein<-DataC$PROPER1>25
high.fat<-DataC$FATPER1>35
low.carb = DataC$CHOPER1<45
low.protien=DataC$PROPER1<15
low.fat = DataC$FATPER1<20

#data<-na.omit(data.frame(BMR=DataC$BMR,Sex=sex,Diet=ifDiet,EIBMR=DataC$EIBMR1))
obese<-factor(ifelse(DataC$BMISC>=30,'1','0'))
y = obese
X = DataC[,c("BDYMSQ04","BMR","SEX","HYPBC","EIBMR1","ADTOTSE")]
summary(X)
summary(y)

```
```{r echo=FALSE}
#Code the categorical variables as factors.
X$SEX = as.numeric(X$SEX=="1")
#sex<-factor(ifelse(DataC$SEX==1,'male','female'))
X$BDYMSQ04=factor(ifelse(X$BDYMSQ04 == 5, 'no diet', 'diet'))
X$HYPBC = as.factor(X$HYPBC)
summary(X)
```

Idenfity the samples containing missing values.
```{r,echo=TRUE}
# This function identifies whether a vector contains any missing values
any.na = function(x) { return(any(is.na(x))); }

# Identify samples with missing value
# ``apply'' the function ``any.na'' to the rows of ``X''
inds1 = which(apply(X,1,any.na))
#Some samples have missing.
inds2 = which(X$BDYMSQ04=="" & X$BMR=="" & X$EIBMR1=="" & X$ADTOTSE=="")
#Cobmine the indicies containng missing values.
inds3 = c(inds1,inds2)
#The following samples contain missing values
length(inds3)
```
There are 170 samples containing missing values and Remove samples that contain missing values
```{r,echo=TRUE}
y = y[-inds3]
X = X[-inds3,]
n = length(y)
```
# LDA

```{r,echo=FALSE}
library(MASS)
# k - value of k in KNN
V =10 #number of CV folds
# seed - (optional) internally sets the seed.
cv.da = function(X,y,method=c("lda","qda"),V,seed=NA)
{
  # Set the seed
  if (!is.na(seed)) {
    set.seed(seed)
  }
  
  # Set n
  n = length(y)
  
  # Split the data up into V folds
  cvSets <- cvFolds(n, V)
  
  # Loop through each fold and calculate the error for that fold
  test.error.da <- c()
  for (i in 1:V) 
  {
    # set the indices corresponding to the training and test sets
    testInds <- cvSets$subsets[which(cvSets$which==i)]
    trainInds <- (1:n)[-testInds]
    
    # Separate y and X into the training and test sets
    y.test <- y[testInds]
    X.test <- X[ testInds,]
    y.train <- y[trainInds]
    X.train <- X[trainInds,]
    
    # Do classification on ith fold
    if (method=="lda") {
      res <- lda(y~., data=X,subset=trainInds)
    }
    if (method=="qda") {
      res <- qda(y~., data=X,subset=trainInds)
    }
    results.da = predict(res, X.test)$class
    
    # Calcuate the test error for this fold
    test.error.da[i] <- sum(results.da!=y.test)
  }
  
  # Calculate the mean error over each fold
  cv.error = sum(test.error.da)/n
  
  # Return the results
  return(cv.error)
}
```

Fit the LDA method
```{r,echo=TRUE}
dat = data.frame(y,X)
X1 = model.matrix(~-1+BMR+SEX+EIBMR1+ADTOTSE,data=dat)
#remove variable +BDYMSQ04+HYPBC?????
X1 = data.frame(X1)
summary(X1)
res <- lda(y~., data=X1,subset=1:n)
res
```
##Comment:
The above output suggests the following interpretations for each of the variables.
* People who have higher BMR are more likely to obese
* Sex = 1 for males. So females are more likely to obese
* Lower EIBMR1(Energy intake) increases obese probability (since the coefficent -3.081292e-01 is negative).
* Lower ADTOTSE(Total mins spent sitting or lying down) increases obese probability (since the coefficent -5.487492e-05is negative).

##CV error
```{r,echo=TRUE}
#cross validation error for LDA.
res.lda = cv.da(X1,y,method="lda",V,seed=1)
res.lda 
```
Comment:
The CV error for LDA is ... percent????


# CART

```{r,echo=TRUE}
library(rpart)
X2 = model.matrix(~-1+BMR+SEX+EIBMR1+ADTOTSE+BDYMSQ04+HYPBC,data=dat)
# Be careful as coding y as a factor here, Otherwise R will do a regression tree rather than a classification tree
res.rpart <- rpart(as.factor(y) ~ X2, data=dat)

library(rpart.plot)
rpart.plot(res.rpart,type=0,extra=1,main="CART fit",cex.main=2,cex=1)
```
##Comment:
Interpret??

cross-validation for rpart.
```{r,echo=FALSE}
cv.rpart = function(X,y,V,seed=NA)
{
  # Set the seed
  if (!is.na(seed)) {
    set.seed(seed)
  }
  
  # Set n
  n = length(y)
  
  # Split the data up into V folds
  cvSets <- cvFolds(n, V)
  
  # Loop through each fold and calculate the error for that fold
  test.error <- c()
  for (i in 1:V) 
  {
    # set the indices corresponding to the training and test sets
    testInds <- cvSets$subsets[which(cvSets$which==i)]
    trainInds <- (1:n)[-testInds]
    
    X = data.frame(X)
    
    # Separate y and X into the training and test sets
    y.test <- y[testInds]
    X.test <- X[ testInds,]
    y.train <- y[trainInds]
    X.train <- X[trainInds,]
  
    # Do classification on ith fold
    res.rpart <- rpart(as.factor(y.train) ~., data=X.train)
    res = predict(res.rpart, newdata=X.test, type = "class")
    
    # Calcuate the test error for this fold
    test.error[i] <- sum(res!=y.test)
  }
  
  # Calculate the mean error over each fold
  cv.error = sum(test.error)/n
  
  # Return the results
  return(cv.error)
}
```

##CV error
```{r,echo=TRUE}
res.rpart = cv.rpart(X2,y,V,seed=1)
res.rpart
```


# Logistic regression

Fit a logistic regression model on most of the data.
```{r,echo=TRUE}
res.glm = glm(y~.,family=binomial,data=X1)
summary(res.glm)
```

The full model has the coefficeints for BMR and SEX as statisticlly
signifficantly different from zero at the 0.05 level. Te fitted model is
$$
\mbox{logit}(p) = -0.1169 + 0.001849 \cdot\mbox{BMR} - 3.258 \cdot \mbox{SEX} - 0.3674 \cdot\mbox{EIBMR1} -0.00006928 \cdot \mbox{ADTOTSE} 
$$
where $p$ is the probability of obesity.

The effect of the significant variables on surival are:
* Larger BMR reduces the probability of obesity.
* Males have reduced the probability of obesity compared to women.



```{r,echo=FALSE}
cv.glm = function(X,y,V,seed=NA)
{
  # Set the seed
  if (!is.na(seed)) {
    set.seed(seed)
  }
  
  # Set n
  n = length(y)
  
  # Split the data up into V folds
  cvSets <- cvFolds(n, V)
  
  # Loop through each fold and calculate the error for that fold
  test.error <- c()
  for (i in 1:V) 
  {
    # set the indices corresponding to the training and test sets
    testInds <- cvSets$subsets[which(cvSets$which==i)]
    trainInds <- (1:n)[-testInds]
    
    X = data.frame(X)
    
    # Separate y and X into the training and test sets
    y.test <- y[testInds]
    X.test <- X[ testInds,]
    y.train <- y[trainInds]
    X.train <- X[trainInds,]
  
    # Do classification on ith fold
    res.glm <- glm(y.train ~., data=X.train, family=binomial)
    res = round(predict(res.glm, newdata=X.test, type="response"))
    
    # Calcuate the test error for this fold
    test.error[i] <- sum(res!=y.test)
  }
  
  # Calculate the mean error over each fold
  cv.error = sum(test.error)/n
  
  # Return the results
  return(cv.error)
}
```

###CV error for glms.
```{r echo=FALSE}
res.glm.cv = cv.glm(X1,y,V,seed=1)
res.glm.cv
```


#summary of cv errors
```{r,echo=FALSE}
ht <- hux(
        Methods = c('LDA',  'rpart', 'glm'),
        Errors = 100*c(min(cv.errors),res.lda,res.rpart,res.glm.cv,), 
        add_colnames = TRUE
      )

bold(ht)[1,]           <- TRUE
bottom_border(ht)[1,]  <- 1
align(ht)[,2]          <- 'right'
right_padding(ht)      <- 10
left_padding(ht)       <- 10
width(ht)              <- 0.35

ht
```




