---
title: "Investigating the relationship between obesity, diet types, and demographics"
author: "Steph Standfort, Lo Ton, and Clara"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{dcolumn}
output:
  html_document:
    highlight: tango
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
always_allow_html: yes
---

```{R, echo = F, include = F, message = F, warning = F}
#setwd("C:/Users/JPRS1/Desktop/STAT3014/Major_project/STAT3014-Major-Project")
proj_dat = read.csv("cleanedData.csv",row.names=1)
dataSteph = proj_dat[which(proj_dat$AGEC >= 18), ]

carb = cut(dataSteph$CHOPER1, breaks=c(-1, 45, 65, 100), labels=c("low", "med", "high"))
protein = cut(dataSteph$PROPER1, breaks=c(-1, 15, 25, 100), labels=c("low", "med", "high"))
fat = cut(dataSteph$FATPER1, breaks=c(-1, 20, 35, 100), labels=c("low", "med", "high"))

library(MASS)
#bmi, age, exercise, 
dat1.var = dataSteph[,c(1,2,8, 11, 12, 14, 77, 86, 64, 65, 68)]
names(dat1.var) = c("bmi", "age", "mins.phys", "waist.cm", "bmr", "ses", "mins.sed", "sex", "protein", "fat", "carbs")

library("kableExtra")
```

#### Predicting Demographics by Diet Type
This section worked out the best diets to model different dependent variables; BMI, BMR, waist size (cm), mintues spent sedentary, sex, SES, and age. The focus of this area was to predict eight key demographics using protein, fat, and carbohydrate diets. Each of the variables was chosen through discussion with NUTM3001 students, and were modelled using linear regression for numerical variables and logistic regression for binary variables. 

Prior to beginning model selection, assumptions for normality of errors and homogeneity of variance were assessed using Q-Q plots and box plots of residuals, and a residual versus fitted plot was used visually assess homogeneity. An example of the graphs used has been provided in Figure 1 showing the diagnostic plots for the dependent variable BMR. For variables BMR and minutes spent sedentary the normality assumption was not met. We decided to apply a log transformation to BMR and a square root transformation to minutes sedentary to satisfy normality. 

```{r, echo = F}
bmr = dat1.var$bmr
bmr.omit = as.data.frame(na.omit(cbind(bmr, protein, fat, carb)))
aovbmr = lm(log(bmr)~protein+fat+carb, data = bmr.omit)
par(mfrow=c(2,2)) 
boxplot(aovbmr$residuals, main = "log(BMR) residuals", ylab = "log(BMR)") 
plot(aovbmr, which=1:3, add.smooth=FALSE)
title("Figure 1", outer=TRUE, line = -1) 
```

Once the the assumptions were met, we began model selection. We used residula deviance tests and AIC Stepwise model selection to find the "best" model for continuous dependent variables and the F test and AIC Stepwise model selection for binary dependent variables. For five of the dependent variables the "best" model was found to be the same using AIC and the residual deviance or F test. For BMR, SES, and minutes spent sedentary the models did not match. When the models did not match, the model selected by the F test was used as the "best" model because the AIC has been claimed to typically prefer overly complex models. 

For each dependent variable using the F or deviance tests, we tested the null hypothesis that the variable was best modelled by the null model (intercept only) compared to the alternative hypothesis that the variable was best modelled by a one factor model. Linear and logistic regression tested the null hypothesis that there is no relationship between the dependent and independent variables compared to the alternative hypothesis that there is a relationship between the dependent variables and atleast one of the independent variables. 
A 5% significance level was chosen as a threshold for the inclusion of the model variables. The resulting "best" models explaining each of the dependent variables is found in Table 1. 

```{r, echo = F}
# Model Selection Table
bmi.mod = rbind("P", "P", "")
bmr.mod = rbind("", "C+PF", "P")
waist.mod = rbind("", "P+C", "P+C")
sed.mod = rbind("", "P+F+C", "P+F+C")
sex.mod = rbind("C+PF", "C+PF","")
ses.mod = rbind("C", "C+PF", "")
age.mod = rbind("", "P+F+C", "P+F+C")

model = t(cbind(bmi.mod, bmr.mod, waist.mod, sed.mod, sex.mod, ses.mod, age.mod))
colnames(model) = c("Deviance", "AIC", "F- test")
rownames(model) = c("BMI", "BMR", "Waist (cm)", "Mins sedentary", "Sex", "SES", "Age")

mod.table = kable_styling(kable(model, booktabs = TRUE, caption ="Table 1: Best models from model selection"),
              latex_options = "hold_position", position = "center")
```

```{r}
mod.table
```

```{r, echo = F, include = F, warning = F, message=F}
bmi.class = cut(dat1.var$bmi, breaks=c(18.5, 30, 65), labels=c("norm","obese")) #categorical
bmi.num = ifelse(bmi.class=="norm", 0,1) #binary

bmi.macro = glm(bmi.num~protein+carb+fat, family = binomial)
best.marco = stepAIC(bmi.macro, scope = list(upper = ~protein*carb*fat, lower = ~1))
summary(best.marco)

#deviance test
mod1 = glm(bmi.num~protein, family = binomial) #signif
mod2 = glm(bmi.num~fat, family = binomial)
mod3 = glm(bmi.num~carb, family = binomial)
a = anova(mod1, test = "Chisq")
b = anova(mod2, test = "Chisq")
c = anova(mod3, test = "Chisq")
a 
b
c

mod4 = glm(bmi.num~protein+fat, family = binomial) #not signif
mod5 = glm(bmi.num~protein+carb, family = binomial) #not signif
d = anova(mod4, test = "Chisq")
e = anova(mod5, test = "Chisq")
d
e

#diet types
pars = coef(best.marco)
logistic = function(x){1/(1+exp(-x))}
prop.low = logistic(pars[1])
prop.med = logistic(pars[1]+pars[2])
prop.high = logistic(pars[1]+pars[2]+pars[3])
prop.low
prop.med
prop.high
```

```{r, echo = F, include = F, warning = F, message = F}
#Waist Assumptions
waist = dat1.var$waist.cm
aovwaist = lm(waist~protein+fat+carb)
par(mfrow=c(2,2)) 
boxplot(aovwaist$residuals) 
plot(aovwaist, which=1:3, add.smooth=FALSE)

#AIC
bestmod.waist = stepAIC(aovwaist , scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.waist)

#deviance test
mod6 = lm(waist~protein)
mod7 = lm(waist~fat)
mod8 = lm(waist~carb)
f = anova(mod6)
g = anova(mod7)
h = anova(mod8)
f #protein is significant and explains the most of the within factor variation out of the three variables. 
g
h 

#protein
mod9 = lm(waist~protein+fat) 
mod10 = lm(waist~protein+carb) #signif
i = anova(mod9)
j = anova(mod10)
i
j

mod11 = lm(waist~protein+carb+fat) #not signif
f = anova(mod11)
f

#mean waist size 
most.ext = 93.3089
mn.high.pro = 93.3089 +1.8730
mn.med.carb = 93.3089 +1.8730 - 0.895

#diet types
prop16 =    93.3089+ 1.8730  #low carb,high prot 
prop17 =  93.3089 + 0.3184 #high fat, med prot 
prop16
prop17

#check final model
waist.chk = lm(waist~protein+carb)
par(mfrow=c(2,2)) 
boxplot(waist.chk$residuals) 
plot(waist.chk, which=1:3, add.smooth=FALSE)
```

```{r, echo = F, warning = F, message = F, include = F}
#BMR
#Assumptions
bmr = dat1.var$bmr
bmr.omit = as.data.frame(na.omit(cbind(bmr, protein, fat, carb)))
aovbmr = lm(log(bmr)~protein+fat+carb, data = bmr.omit)
par(mfrow=c(2,2)) 
boxplot(aovbmr$residuals) 
plot(aovbmr, which=1:3, add.smooth=FALSE)

#AIC
bestmod.bmr = stepAIC(aovbmr , scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.bmr)

#F test
mod12 = lm(log(bmr)~protein, data = bmr.omit) 
mod13 = lm(log(bmr)~fat, data = bmr.omit)
mod14 = lm(log(bmr)~carb, data = bmr.omit) #most signif at 0.1 level
k = anova(mod12)
l = anova(mod13)
m = anova(mod14)
k  
l
m

#proportion
prop13 =   6792.647 +  24.462+ 47.487#low carb, med fat, high prot 
prop14 =  6792.647 -220.555 -112.766  #high fat, lwo carb, med prot 
prop15 =  6792.647 -190.932 -112.766 #high carb, med prot, low fat 
prop13
prop14
prop15

#check assumptions of chosen model
bmr.chk = lm(log(bmr)~protein + fat + carb + protein:fat + protein:carb + fat:carb, data = bmr.omit)
par(mfrow=c(2,2)) 
boxplot(bmr.chk$residuals) 
plot(bmr.chk, which=1:3, add.smooth=FALSE)
```

```{r, echo = F, include = F, warning = F, message = F}
#Sedentary
#Assumptions for linear regression
sed = dat1.var$mins.sed
sed.omit = as.data.frame(na.omit(cbind(sed, protein, fat, carb)))
aovsed = lm(sqrt(sed)~protein+fat+carb, data = sed.omit)
par(mfrow=c(2,2)) 
boxplot(aovsed$residuals) 
plot(aovsed, which=1:3, add.smooth=FALSE)

#AIC
bestmod.sed = stepAIC(aovsed , scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.sed)

#deviance test
mod15 = lm(sqrt(sed)~protein, data = sed.omit)
mod16 = lm(sqrt(sed)~fat, data = sed.omit)
mod17 = lm(sqrt(sed)~carb, data = sed.omit)
n = anova(mod15)
o = anova(mod16)
p = anova(mod17)
n 
o #most signif
p

#fat
mod18 = lm(sqrt(sed)~fat+protein, data = sed.omit) #signif
mod19 = lm(sqrt(sed)~fat+carb, data = sed.omit)
q = anova(mod18)
r = anova(mod19)
q
r

#fat+protein
mod20 = lm(sqrt(sed)~fat+protein+carb, data = sed.omit)#signif
s = anova(mod20)
s 

#interaction
mod21 = lm(sqrt(sed)~fat*protein*carb, data = sed.omit) #not signif
t = anova(mod21)
t 

#mean of each group
prop10 =  2350.15 + 108.18-198.91#low carb, med fat, high prot 37hrs
prop11 =  2350.15+ 131.18 -92.92  #high fat, lwo carb, med prot 39hrs
prop12 = 2350.15 -92.92 -203.00#high carb, med prot, low fat #34hrs
prop10 #atkin
prop11 #keto
prop12 #dash

#Check best model - same as before
```

```{R, echo = F, include = F, warning = F, message = F}
#Sex
#AIC
sex = ifelse(dat1.var$sex==1,1,0) #1 = male
glm.sex = glm(sex~protein+fat+carb, family= binomial)
bestmod.sex = stepAIC(glm.sex, scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.sex)

#deviance test
mod22 = lm(sex~protein, family = "binomial")
mod23 = lm(sex~fat,  family = "binomial")
mod24 = lm(sex~carb,  family = "binomial")
u = anova(mod22, test = "Chisq")
v = anova(mod23, test = "Chisq")
w= anova(mod24, test = "Chisq")
u 
v #signif: explains more within SS variation
w  #signif

#fat
mod25 = glm(sex~fat+protein, family = "binomial") #not signif
mod26 = glm(sex~fat+carb, family = "binomial") #signif
x = anova(mod25, test = "Chisq")
y = anova(mod26, test = "Chisq")
x
y

#fat+carb
mod27 = glm(sex~fat+carb+protein, family = "binomial") #signif
z = anova(mod27, test = "Chisq") 
z

fat[which.max(glm.sex$fitted.values)]
protein[which.max(glm.sex$fitted.values)]
carb[which.max(glm.sex$fitted.values)]
max(glm.sex$fitted.values)
prop = logistic(0.46303-0.31307-0.48824) #low carb, med fat, high protein
pro0 = logistic(0.46303 -0.25332-0.71966)#high carb, low fat, med protein
prop1 = logistic( 0.46303  -0.27346-0.25332)#med carb, low fat, med prot
prop2 = logistic( 0.46303 -0.68138-0.25332)#high fat, low carb, med prot
prop
pro0
prop1
prop2

```

```{r, echo = F, include = F, warning = F, message = F}
#SES
#AIC
dataSteph$SF2SA1QN = as.numeric(as.factor(dataSteph$SF2SA1QN))
ses = ifelse(dataSteph$SF2SA1QN>4,1,0) #decile [1:4] = 0
glmadd.ses = glm(ses~protein+fat+carb, family= binomial)
bestmod.ses = stepAIC(glmadd.ses, scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.ses)

#Deviance
mod28 = glm(ses~protein, family = "binomial")
mod29 = glm(ses~fat,  family = "binomial")
mod30 = glm(ses~carb,  family = "binomial") #signif
aa = anova(mod28, test = "Chisq")
bb = anova(mod29, test = "Chisq")
cc = anova(mod30, test = "Chisq")
aa 
bb
cc 

#carbs
mod31 = glm(ses~carb+protein, family = "binomial") #not signif
mod32 = glm(ses~carb+fat, family = "binomial") #not signif
dd = anova(mod31, test = "Chisq")
ee = anova(mod32, test = "Chisq")
dd
ee

#percentages
fat[which.max(glmadd.ses$fitted.values)]
protein[which.max(glmadd.ses$fitted.values)]
carb[which.max(glmadd.ses$fitted.values)]
max(glmadd.ses$fitted.values)

#diet types
prop3 = logistic( -1.18602+0.13784-0.12254) #low carb, med fat, high protein
prop4 = logistic(-1.18602-0.20883+0.31338)#med carb, low fat, med prot
prop5 = logistic(-1.18602 -0.04306+ 0.31338)#high fat, low carb, med prot
prop6 = logistic(-1.18602 -0.53563+0.31338)#high carb, low fat, med prot
prop3
prop4
prop5
prop6
```

```{r,  echo = F, include = F, warning = F, message = F}
#Age
#Assumptions for linear regression
age = dataSteph$AGEC
aovadd.age = lm(age~protein+fat+carb)
par(mfrow=c(2,2)) 
boxplot(aovadd.age$residuals) 
plot(aovadd.age, which=1:3, add.smooth=FALSE)

#AIC
bestmod.age = stepAIC(aovadd.age, scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.age)

mod33 = lm(age~protein) #signif
mod34 = lm(age~fat)
mod35 = lm(age~carb)
ff = anova(mod33)
gg = anova(mod34)
hh = anova(mod35)
ff  
gg 
hh

#protein
mod36 = lm(age~protein+carb) #signif
mod37 = lm(age~protein+fat)
ii = anova(mod36)
jj = anova(mod37)
ii
jj

mod38 = lm(age~protein+carb+fat) #signif
kk = anova(mod38)
kk


mod39 = lm(age~protein*carb*fat) #not signif
ll = anova(mod39)
ll

#expected mean
fat[which.max(aovadd.age$fitted.values)]
protein[which.max(aovadd.age$fitted.values)]
carb[which.max(aovadd.age$fitted.values)]
max(aovadd.age$fitted.values)

#diet types
prop7 = 51.2510-0.7956-0.4439      #low carb, med fat, high protein Y
prop8 =  51.2510 -3.1630 + 0.7247  #high fat, low carb, med prot Y
prop9 =  51.2510-6.2463+0.7247#high carb, low fat, med prot Y
prop7
prop8
prop9

#check selected model- same as above
```

# Demographic Table

```{R, echo = F, include = F, message = F, warning = F}
atkin= dat1.var[which(dat1.var$protein>20 & dat1.var$carbs<45 & dat1.var$fat>20 & dat1.var$fat<35),] 
keto = dat1.var[which(dat1.var$fat>35 & dat1.var$carbs<45 & dat1.var$protein>15 & dat1.var$protein<25),]
dash = dat1.var[which(dat1.var$fat<20 & dat1.var$carbs>65 & dat1.var$protein>15 & dat1.var$protein<25),]

#atkin (high pro, low carb), keti (high fat, low carb), dash(high carb, low fat)
bmi.row = rbind(mean(na.omit(atkin$bmi)), mean(na.omit(keto$bmi)), mean(na.omit(dash$bmi)))
bmr.row = rbind(mean(na.omit(atkin$bmr)), mean(na.omit(keto$bmr)), mean(na.omit(dash$bmr)))
waist.row = rbind(mean(na.omit(atkin$waist)), mean(na.omit(keto$waist)), mean(na.omit(dash$waist)))
sed.row = rbind(prop10, prop11, prop12) 
sex.row = rbind(prop, prop2, pro0)
ses.row = rbind(prop3, prop5, prop6)
age.row = rbind(mean(na.omit(atkin$age)), mean(na.omit(keto$age)), mean(na.omit(dash$age)))

result = t(cbind(bmi.row, bmr.row, waist.row, sed.row, sex.row, ses.row, age.row))
colnames(result) = c("Atkins", "Keto", "Low Fat, High Carb")
rownames(result) = c("BMI", "BMR", "Waist (cm)", "Mins sedentary", "Sex", "SES", "Age")
result = round(result, digits = 3)

```

#### Demographics for diets

Table 2 presents the key characteristics of each of the three diets investigated. For BMR, waist measurement, minutes spent sedentary, and age, the mean of each group was reported. For BMI, sex, and SES respectively, the probability of a person being overweight, male, or in the normal decile Index of Relative Socio-Economic Disadvantage has been reported. 

```{r, echo = F}
#demographic table
result.tab = kable_styling(kable(result, booktabs = TRUE, caption ="Table 2: Demographics for 8 key variables"),
              latex_options = "hold_position", position = "center")
```

```{r}
result.tab
```

##### A couple of key things to note:

* BMI: The probability of being obese when a person has a high protein diet in this sample is 40.1%. A possible explanation of this may be that people on high protein diets could be focussed on muscle development which weights more than fat, resulting in a higher body weight, and consequently a higher BMI. 

* Waist: The largest mean waist size in the dietary model of waist size was 95.12cm for people with a high protein diet. 
* Sedentary: The high fat low carbohyrate diet (keto diet) had the highest mean sedentary minutes of 2388.410, which equates to 39.8 hours of being sedentary over two days. 

* Sex: The expected proportion of males in the the low fat, low protein, low carb group was 59%. The expected proportion of males in the low carb, med fat and high protein is 63% 

* SES: In the low fat, medium protein, low carb diet a person from medium SES will be in this diet with a probability of 26%. 

* Age: The most extreme value is in the low fat, medium protein, low carb group. The mean age in this group is 52 years. 






