---
output:
  html_document: default
  pdf_document: default
---
##Major Project Data Exploration
BMI, or body mass index, is a number calculated from your height and weight which is then used to assess your body composition. Your BMR, or basal metabolic rate, is the number of calories you burn when your body is at rest.
Question: dieting status (BDYMSQ04) - 0 corresponds to NA. Does this mean its a missing value or does it mean participant is not on a diet. Depending on what 0 is will change the possible meaning of results. 
```{R, include = FALSE}
dat_orig = read.csv("C:/Users/JPRS1/Desktop/Datasets/nutmstatData2018.csv")
#dat_orig = read.csv("/users/ugrad/ssta9076/Desktop/nutmstatData2018.csv")
#install.packages("huxtable")
#install.packages("factoextra")
#library(factoextra)
#library(huxtable )
#library(ggplot2)
#install.packages("qwraps2")
library(qwraps2)
#dim(dat_orig)
library(dplyr)

dat <-  dat_orig

colNames <- colnames(dat_orig)

# for (i in 1:length(colNames)) { cat("categoricalList[",i,"] <- FALSE # ",colNames[i],"\n"); }
#categorical or numerical

categoricalList <- c()
categoricalList[ 1 ] <- FALSE #  BMISC 
categoricalList[ 2 ] <- FALSE #  AGEC 
categoricalList[ 3 ] <- TRUE #  SMSBC 
categoricalList[ 4 ] <- TRUE #  COBBC 
categoricalList[ 5 ] <- TRUE #  FEMLSBC 
categoricalList[ 6 ] <- FALSE #  PHDKGWBC 
categoricalList[ 7 ] <- FALSE #  PHDCMHBC 
categoricalList[ 8 ] <- FALSE #  EXLWTBC 
categoricalList[ 9 ] <- FALSE #  EXLWMBC 
categoricalList[ 10 ] <- FALSE #  EXLWVBC 
categoricalList[ 11 ] <- FALSE #  PHDCMWBC 
categoricalList[ 12 ] <- FALSE #  BMR 
categoricalList[ 13 ] <- FALSE #  EIBMR1 
categoricalList[ 14 ] <- TRUE #  SF2SA1QN 
categoricalList[ 15 ] <- TRUE #  INCDEC 
categoricalList[ 16 ] <- TRUE #  DIABBC 
categoricalList[ 17 ] <- TRUE #  HCHOLBC 
categoricalList[ 18 ] <- TRUE #  HSUGBC 
categoricalList[ 19 ] <- TRUE #  HYPBC 
categoricalList[ 20 ] <- FALSE #  ENERGYT1 
categoricalList[ 21 ] <- FALSE #  ENRGYT1 
categoricalList[ 22 ] <- FALSE #  MOISTT1 
categoricalList[ 23 ] <- FALSE #  PROTT1 
categoricalList[ 24 ] <- FALSE #  FATT1 
categoricalList[ 25 ] <- FALSE #  CHOWSAT1 
categoricalList[ 26 ] <- FALSE #  CHOWOAT1 
categoricalList[ 27 ] <- FALSE #  STARCHT1 
categoricalList[ 28 ] <- FALSE #  SUGART1 
categoricalList[ 29 ] <- FALSE #  FIBRET1 
categoricalList[ 30 ] <- FALSE #  ALCT1 
categoricalList[ 31 ] <- FALSE #  PREVAT1 
categoricalList[ 32 ] <- FALSE #  PROVAT1 
categoricalList[ 33 ] <- FALSE #  RETEQT1 
categoricalList[ 34 ] <- FALSE #  B1T1 
categoricalList[ 35 ] <- FALSE #  B2T1 
categoricalList[ 36 ] <- FALSE #  B3T1 
categoricalList[ 37 ] <- FALSE #  NIACINT1 
categoricalList[ 38 ] <- FALSE #  FOLATT1 
categoricalList[ 39 ] <- FALSE #  FOLICT1 
categoricalList[ 40 ] <- FALSE #  TOTFOLT1 
categoricalList[ 41 ] <- FALSE #  FOLEQT1 
categoricalList[ 42 ] <- FALSE #  B6T1 
categoricalList[ 43 ] <- FALSE #  B12T1 
categoricalList[ 44 ] <- FALSE #  VITCT1 
categoricalList[ 45 ] <- FALSE #  VITET1 
categoricalList[ 46 ] <- FALSE #  CALCT1 
categoricalList[ 47 ] <- FALSE #  IODINET1 
categoricalList[ 48 ] <- FALSE #  IRONT1 
categoricalList[ 49 ] <- FALSE #  MAGT1 
categoricalList[ 50 ] <- FALSE #  PHOST1 
categoricalList[ 51 ] <- FALSE #  POTAST1 
categoricalList[ 52 ] <- FALSE #  SELT1 
categoricalList[ 53 ] <- FALSE #  SODIUMT1 
categoricalList[ 54 ] <- FALSE #  ZINCT1 
categoricalList[ 55 ] <- FALSE #  CAFFT1 
categoricalList[ 56 ] <- FALSE #  CHOLT1 
categoricalList[ 57 ] <- FALSE #  SATFATT1 
categoricalList[ 58 ] <- FALSE #  MUFATT1 
categoricalList[ 59 ] <- FALSE #  PUFATT1 
categoricalList[ 60 ] <- FALSE #  LAT1 
categoricalList[ 61 ] <- FALSE #  ALAT1 
categoricalList[ 62 ] <- FALSE #  LCN3T1 
categoricalList[ 63 ] <- FALSE #  TRANST1 
categoricalList[ 64 ] <- FALSE #  PROPER1 
categoricalList[ 65 ] <- FALSE #  FATPER1 
categoricalList[ 66 ] <- FALSE #  LAPER1 
categoricalList[ 67 ] <- FALSE #  ALAPER1 
categoricalList[ 68 ] <- FALSE #  CHOPER1 
categoricalList[ 69 ] <- FALSE #  SUGPER1 
categoricalList[ 70 ] <- FALSE #  STARPER1 
categoricalList[ 71 ] <- FALSE #  ALCPER1 
categoricalList[ 72 ] <- FALSE #  SATPER1 
categoricalList[ 73 ] <- FALSE #  TRANPER1 
categoricalList[ 74 ] <- FALSE #  FIBRPER1 
categoricalList[ 75 ] <- FALSE #  MONOPER1 
categoricalList[ 76 ] <- FALSE #  POLYPER1 
categoricalList[ 77 ] <- FALSE #  ADTOTSE 
categoricalList[ 78 ] <- TRUE #  BDYMSQ04 given as NAs values by John 
categoricalList[ 79 ] <- FALSE #  DIASTOL 
categoricalList[ 80 ] <- TRUE #  DIETQ12 
categoricalList[ 81 ] <- TRUE #  DIETQ14 
categoricalList[ 82 ] <- TRUE #  DIETQ5 
categoricalList[ 83 ] <- TRUE #  DIETQ8 
categoricalList[ 84 ] <- TRUE #  DIETRDI 
categoricalList[ 85 ] <- TRUE #  SABDYMS 
categoricalList[ 86 ] <- TRUE #  SEX 
categoricalList[ 87 ] <- FALSE #  SLPTIME 
categoricalList[ 88 ] <- TRUE #  SMKDAILY 
categoricalList[ 89 ] <- TRUE #  SMKSTAT 
categoricalList[ 90 ] <- FALSE #  SYSTOL 
categoricalList[ 91 ] <- TRUE #  ABSPID 
categoricalList[ 92 ] <- FALSE #  GRAINS1N 
categoricalList[ 93 ] <- FALSE #  WHOLGR1N 
categoricalList[ 94 ] <- FALSE #  REFGRA1N 
categoricalList[ 95 ] <- FALSE #  VEGLEG1N 
categoricalList[ 96 ] <- FALSE #  GREENS1N 
categoricalList[ 97 ] <- FALSE #  VGORSV1N 
categoricalList[ 98 ] <- FALSE #  STARCH1N 
categoricalList[ 99 ] <- FALSE #  LEGVEG1N 
categoricalList[ 100 ] <- FALSE #  OTHVEG1N 
categoricalList[ 101 ] <- FALSE #  FRUIT1N 
categoricalList[ 102 ] <- FALSE #  FRJUIC1N 
categoricalList[ 103 ] <- FALSE #  DAIRY1N 
categoricalList[ 104 ] <- FALSE #  MEAT1N 
categoricalList[ 105 ] <- FALSE #  RDMTL1N 
categoricalList[ 106 ] <- FALSE #  RDMTLU1N 
categoricalList[ 107 ] <- FALSE #  RDMTN1N 
categoricalList[ 108 ] <- FALSE #  RDMTNU1N 
categoricalList[ 109 ] <- FALSE #  RDMTNP1N 
categoricalList[ 110 ] <- FALSE #  PLTYL1N 
categoricalList[ 111 ] <- FALSE #  PLTYLU1N 
categoricalList[ 112 ] <- FALSE #  FISH1N 
categoricalList[ 113 ] <- FALSE #  EGGS1N 
categoricalList[ 114 ] <- FALSE #  LEGMT1N 
categoricalList[ 115 ] <- FALSE #  NUTS1N 
categoricalList[ 116 ] <- FALSE #  WATER1N 
categoricalList[ 117 ] <- FALSE #  UNSAT1N 
categoricalList[ 118 ] <- FALSE #  FRESUG1N 
categoricalList[ 119 ] <- FALSE #  ADDSUG1N 
categoricalList[ 120 ] <- FALSE #  WATERG1N 
categoricalList[ 121 ] <- FALSE #  PEFRESD1 
categoricalList[ 122 ] <- FALSE #  PEADDSD1 
categoricalList[ 123 ] <- TRUE #  ALTNTR 
categoricalList[ 124 ] <- TRUE #  ALTRESB 
categoricalList[ 125 ] <- TRUE #  APOBNTR 
categoricalList[ 126 ] <- TRUE #  APOBRESB 
categoricalList[ 127 ] <- TRUE #  B12RESB 
categoricalList[ 128 ] <- TRUE #  BIORESPC 
categoricalList[ 129 ] <- TRUE #  CHOLNTR 
categoricalList[ 130 ] <- TRUE #  CHOLRESB 
categoricalList[ 131 ] <- TRUE #  CVDMEDST 
categoricalList[ 132 ] <- TRUE #  DIAHBRSK 
categoricalList[ 133 ] <- TRUE #  FASTSTAD 
categoricalList[ 134 ] <- TRUE #  FOLATREB 
categoricalList[ 135 ] <- TRUE #  GGTNTR 
categoricalList[ 136 ] <- TRUE #  GGTRESB 
categoricalList[ 137 ] <- TRUE #  GLUCFPD 
categoricalList[ 138 ] <- TRUE #  GLUCFREB 
categoricalList[ 139 ] <- TRUE #  HBA1PREB 
categoricalList[ 140 ] <- TRUE #  HDLCHREB 
categoricalList[ 141 ] <- TRUE #  LDLNTR 
categoricalList[ 142 ] <- TRUE #  LDLRESB 
categoricalList[ 143 ] <- TRUE #  TRIGNTR 
categoricalList[ 144 ] <- TRUE #  TRIGRESB 

for (i in 1:length(categoricalList)) {
  if (categoricalList[ i ]) {
      dat[,i] <- as.factor(dat[  ,i])
  }
}




# for (i in 1:length(colNames)) { cat("excludeList[[",i,"]] <- c() # ",colNames[i],"\n"); }
#removes certain observations

excludeList <- list()

excludeList[[ 1 ]] <- c(0,98,99) #  BMISC 
excludeList[[ 2 ]] <- c() #  AGEC 
excludeList[[ 3 ]] <- c(0) #  SMSBC 
excludeList[[ 4 ]] <- c() #  COBBC 
excludeList[[ 5 ]] <- c() #  FEMLSBC 
excludeList[[ 6 ]] <- c(0,997,998,999) #  PHDKGWBC 
excludeList[[ 7 ]] <- c(0,998,999) #  PHDCMHBC 
excludeList[[ 8 ]] <- c(9996, 9999) #  EXLWTBC 
excludeList[[ 9 ]] <- c() #  EXLWMBC 
excludeList[[ 10 ]] <- c() #  EXLWVBC 
excludeList[[ 11 ]] <- c(0,998,999) #  PHDCMWBC 
excludeList[[ 12 ]] <- c(99998) #  BMR 
excludeList[[ 13 ]] <- c(998) #  EIBMR1 
excludeList[[ 14 ]] <- c(0,99) #  SF2SA1QN 
excludeList[[ 15 ]] <- c(0,98,99) #  INCDEC 
excludeList[[ 16 ]] <- c() #  DIABBC 
excludeList[[ 17 ]] <- c() #  HCHOLBC 
excludeList[[ 18 ]] <- c() #  HSUGBC 
excludeList[[ 19 ]] <- c() #  HYPBC 
excludeList[[ 20 ]] <- c() #  ENERGYT1 
excludeList[[ 21 ]] <- c() #  ENRGYT1 
excludeList[[ 22 ]] <- c() #  MOISTT1 
excludeList[[ 23 ]] <- c() #  PROTT1 
excludeList[[ 24 ]] <- c() #  FATT1 
excludeList[[ 25 ]] <- c() #  CHOWSAT1 
excludeList[[ 26 ]] <- c() #  CHOWOAT1 
excludeList[[ 27 ]] <- c() #  STARCHT1 
excludeList[[ 28 ]] <- c() #  SUGART1 
excludeList[[ 29 ]] <- c() #  FIBRET1 
excludeList[[ 30 ]] <- c() #  ALCT1 
excludeList[[ 31 ]] <- c() #  PREVAT1 
excludeList[[ 32 ]] <- c() #  PROVAT1 
excludeList[[ 33 ]] <- c() #  RETEQT1 
excludeList[[ 34 ]] <- c() #  B1T1 
excludeList[[ 35 ]] <- c() #  B2T1 
excludeList[[ 36 ]] <- c() #  B3T1 
excludeList[[ 37 ]] <- c() #  NIACINT1 
excludeList[[ 38 ]] <- c() #  FOLATT1 
excludeList[[ 39 ]] <- c() #  FOLICT1 
excludeList[[ 40 ]] <- c() #  TOTFOLT1 
excludeList[[ 41 ]] <- c() #  FOLEQT1 
excludeList[[ 42 ]] <- c() #  B6T1 
excludeList[[ 43 ]] <- c() #  B12T1 
excludeList[[ 44 ]] <- c(10144.97) #  VITCT1 
excludeList[[ 45 ]] <- c() #  VITET1 
excludeList[[ 46 ]] <- c() #  CALCT1 
excludeList[[ 47 ]] <- c() #  IODINET1 
excludeList[[ 48 ]] <- c() #  IRONT1 
excludeList[[ 49 ]] <- c() #  MAGT1 
excludeList[[ 50 ]] <- c() #  PHOST1 
excludeList[[ 51 ]] <- c() #  POTAST1 
excludeList[[ 52 ]] <- c() #  SELT1 
excludeList[[ 53 ]] <- c() #  SODIUMT1 
excludeList[[ 54 ]] <- c() #  ZINCT1 
excludeList[[ 55 ]] <- c() #  CAFFT1 
excludeList[[ 56 ]] <- c() #  CHOLT1 
excludeList[[ 57 ]] <- c() #  SATFATT1 
excludeList[[ 58 ]] <- c() #  MUFATT1 
excludeList[[ 59 ]] <- c() #  PUFATT1 
excludeList[[ 60 ]] <- c() #  LAT1 
excludeList[[ 61 ]] <- c() #  ALAT1 
excludeList[[ 62 ]] <- c() #  LCN3T1 
excludeList[[ 63 ]] <- c() #  TRANST1 
excludeList[[ 64 ]] <- c() #  PROPER1 
excludeList[[ 65 ]] <- c() #  FATPER1 
excludeList[[ 66 ]] <- c() #  LAPER1 
excludeList[[ 67 ]] <- c() #  ALAPER1 
excludeList[[ 68 ]] <- c() #  CHOPER1 
excludeList[[ 69 ]] <- c() #  SUGPER1 
excludeList[[ 70 ]] <- c() #  STARPER1 
excludeList[[ 71 ]] <- c() #  ALCPER1 
excludeList[[ 72 ]] <- c() #  SATPER1 
excludeList[[ 73 ]] <- c() #  TRANPER1 
excludeList[[ 74 ]] <- c() #  FIBRPER1 
excludeList[[ 75 ]] <- c() #  MONOPER1 
excludeList[[ 76 ]] <- c() #  POLYPER1 
excludeList[[ 77 ]] <- c(0, 9999, 9996) #  ADTOTSE 
excludeList[[ 78 ]] <- c(0,6) #  BDYMSQ04 
excludeList[[ 79 ]] <- c(0,998,999) #  DIASTOL 
excludeList[[ 80 ]] <- c(0,6) #  DIETQ12 
excludeList[[ 81 ]] <- c(0,6) #  DIETQ14 
excludeList[[ 82 ]] <- c(0) #  DIETQ5 
excludeList[[ 83 ]] <- c(0) #  DIETQ8 
excludeList[[ 84 ]] <- c(0,3) #  DIETRDI 
excludeList[[ 85 ]] <- c(0,8,9) #  SABDYMS 
excludeList[[ 86 ]] <- c() #  SEX 
excludeList[[ 87 ]] <- c(0,9998,9999) #  SLPTIME 
excludeList[[ 88 ]] <- c() #  SMKDAILY 
excludeList[[ 89 ]] <- c(0) #  SMKSTAT 
excludeList[[ 90 ]] <- c(0,998,999) #  SYSTOL 
excludeList[[ 91 ]] <- c() #  ABSPID 
excludeList[[ 92 ]] <- c() #  GRAINS1N 
excludeList[[ 93 ]] <- c() #  WHOLGR1N 
excludeList[[ 94 ]] <- c() #  REFGRA1N 
excludeList[[ 95 ]] <- c() #  VEGLEG1N 
excludeList[[ 96 ]] <- c() #  GREENS1N 
excludeList[[ 97 ]] <- c() #  VGORSV1N 
excludeList[[ 98 ]] <- c() #  STARCH1N 
excludeList[[ 99 ]] <- c() #  LEGVEG1N 
excludeList[[ 100 ]] <- c() #  OTHVEG1N 
excludeList[[ 101 ]] <- c() #  FRUIT1N 
excludeList[[ 102 ]] <- c() #  FRJUIC1N 
excludeList[[ 103 ]] <- c() #  DAIRY1N 
excludeList[[ 104 ]] <- c() #  MEAT1N 
excludeList[[ 105 ]] <- c() #  RDMTL1N 
excludeList[[ 106 ]] <- c() #  RDMTLU1N 
excludeList[[ 107 ]] <- c() #  RDMTN1N 
excludeList[[ 108 ]] <- c() #  RDMTNU1N 
excludeList[[ 109 ]] <- c() #  RDMTNP1N 
excludeList[[ 110 ]] <- c() #  PLTYL1N 
excludeList[[ 111 ]] <- c() #  PLTYLU1N 
excludeList[[ 112 ]] <- c() #  FISH1N 
excludeList[[ 113 ]] <- c() #  EGGS1N 
excludeList[[ 114 ]] <- c() #  LEGMT1N 
excludeList[[ 115 ]] <- c() #  NUTS1N 
excludeList[[ 116 ]] <- c() #  WATER1N 
excludeList[[ 117 ]] <- c() #  UNSAT1N 
excludeList[[ 118 ]] <- c() #  FRESUG1N 
excludeList[[ 119 ]] <- c() #  ADDSUG1N 
excludeList[[ 120 ]] <- c() #  WATERG1N 
excludeList[[ 121 ]] <- c() #  PEFRESD1 
excludeList[[ 122 ]] <- c() #  PEADDSD1 
excludeList[[ 123 ]] <- c(0,8) #  ALTNTR 
excludeList[[ 124 ]] <- c(97,98) #  ALTRESB 
excludeList[[ 125 ]] <- c(0,8) #  APOBNTR 
excludeList[[ 126 ]] <- c(97,98) #  APOBRESB 
excludeList[[ 127 ]] <- c(97,98) #  B12RESB 
excludeList[[ 128 ]] <- c(0) #  BIORESPC 
excludeList[[ 129 ]] <- c(0,8) #  CHOLNTR 
excludeList[[ 130 ]] <- c(97,98) #  CHOLRESB 
excludeList[[ 131 ]] <- c(0,8) #  CVDMEDST 
excludeList[[ 132 ]] <- c(0,8) #  DIAHBRSK 
excludeList[[ 133 ]] <- c(0) #  FASTSTAD 
excludeList[[ 134 ]] <- c(97,98) #  FOLATREB 
excludeList[[ 135 ]] <- c(8) #  GGTNTR 
excludeList[[ 136 ]] <- c(97,98) #  GGTRESB 
excludeList[[ 137 ]] <- c(0,8) #  GLUCFPD 
excludeList[[ 138 ]] <- c(97,98) #  GLUCFREB 
excludeList[[ 139 ]] <- c(7,8) #  HBA1PREB 
excludeList[[ 140 ]] <- c(7,8) #  HDLCHREB 
excludeList[[ 141 ]] <- c(0,8) #  LDLNTR 
excludeList[[ 142 ]] <- c(97,98) #  LDLRESB 
excludeList[[ 143 ]] <- c(0,8) #  TRIGNTR 
excludeList[[ 144 ]] <- c(97,98) #  TRIGRESB 

for (i in 1:length(excludeList)) {
  dat[ dat[,i]%in%excludeList[[ i ]],i] <- NA
}

```

```{R, include = F}
#keep variables with less than 50% missing values
keepR = which(apply(is.na(dat), 1, mean)<= 1) #keep rows where there are less than 100% NAs
keepC = which(apply(is.na(dat), 2, mean) <= 0.5) #keep columns where there are less than 50% NAs
proj_dat = dat[keepR, keepC]
sum(is.na(dat$GLUCFPD))
```


##High protein, low carb diet, no children, split into diet and no diet
```{r}
library(plyr)
proj_dat_diet2 = proj_dat[which(proj_dat$PROPER1 > 20 & proj_dat$CHOPER1 < 45 & proj_dat$AGEC >= 18), ]
proj_dat_diet2$BDYMSQ04 = mapvalues(proj_dat_diet2$BDYMSQ04, from = c("1", "2", "3", "5"), to = c("8", "8", "8", "9")) #8 = diet, 9 = no diet
```

####BMI and BMR in high protien and low carb diet
```{r}
#have included separation by diet type but it may not be too important to do this at this stage
qplot(BMISC, BMR, colour = BDYMSQ04, data = proj_dat_diet2) + labs(title = "BMI and BMR, diet") #DIET TYPE
qplot(BMISC, BMR, colour = SEX, data = proj_dat_diet2) + labs(title = "BMI and BMR, gender") #SEX
qplot(BMISC, EIBMR1, colour = BDYMSQ04, data = proj_dat_diet2) + labs(title = "BMI and energy intake, diet")
qplot(BMISC, EIBMR1, colour = SEX, data = proj_dat_diet2) + labs(title = "BMI and energy intake, diet")
qplot(BMR, EIBMR1, colour = BDYMSQ04, data = proj_dat_diet2) + labs(title = "BMI and energy intake, gender") 
qplot(BMR, EIBMR1, colour = SEX, data = proj_dat_diet2) + labs(title = "BMI and energy intake, gender") 

ggplot(data = proj_dat_diet2, aes(x=BDYMSQ04, y = BMISC, fill = BDYMSQ04)) + geom_boxplot()
ggplot(data = proj_dat_diet2, aes(x=BDYMSQ04, y = BMR, fill = SEX)) + geom_boxplot()
ggplot(data = proj_dat_diet2, aes(x = SEX, y = BMISC, fill = BDYMSQ04)) + geom_boxplot()
ggplot(data = proj_dat_diet2, aes(x = SEX, y = BMR, fill = BDYMSQ04)) + geom_boxplot()
ggplot(data = proj_dat_diet2, aes(x = BDYMSQ04, y = BMISC)) + geom_boxplot()
```

Comment: cant see a different in BMI/ BMR across different genders or across dieting and not dieting in adults. Can see that those dieting have a higher BMI across both genders than those who are not dieting. Can also see than males have higher BMRs than females for both dieting and not dieting. In both gender those who are dieting have higher BMRs than those not dieting.  

```{r}
mean.rm.na <- function(x) mean(x,na.rm=T)
interaction.plot(x.factor = proj_dat_diet2$BDYMSQ04, trace.factor = proj_dat_diet2$SEX, response = proj_dat_diet2$BMISC, fun = mean.rm.na, main="Interaction between dieting status across gender", xlab = "Diet or no diet", ylab = "Mean BMI")

interaction.plot(x.factor = proj_dat_diet2$SEX, trace.factor = proj_dat_diet2$BDYMSQ04, response = proj_dat_diet2$BMISC, fun = mean.rm.na, main="Interaction between dieting status across gender", xlab = "Diet or no diet", ylab = "Mean BMI")

# legend("topright", c("Male","Female"),bty="n",lty=c(1,2),lwd=2,pch=c(18,24), col=c("red","blue"), title="Sex",inset = .02)
```
Comment: no interaction between genders across diet but there is an interaction between diet type across gender. Worthy to do an ANOVA?

####Varible selection
```{r}
install.packages("bestglm")
library(bestglm)

data = proj_dat_diet2[,c("BMISC", "BDYMSQ04","BMR","SEX","HYPBC","EIBMR1","ADTOTSE")]
y1 = data$BMISC
X1= data[c("BDYMSQ04","BMR","SEX","HYPBC","EIBMR1","ADTOTSE")]
xy = cbind(X1,y1)
glm1 = bestglm(xy,IC = "BIC",method = "forward") #NAS
glm1$Subsets

glm1.1 = bestglm(na.omit(xy),IC = "BIC",method = "forward") #no NAS
as.logical(glm1$Subsets[1,2:(ncol(glm1$Subsets)-2)])
X1[,as.logical(glm1$Subsets[2,2:(ncol(glm1$Subsets)-2)])]
glm1$Subsets[2,2:(ncol(glm1$Subsets)-2)]
CVHTF(X1[,as.logical(glm1$Subsets[3,2:(ncol(glm1$Subsets)-2)])],y1,K=10)
install.packages("DAAG")
library(DAAG)
??cvTools

(nrow(na.omit(xy)))/(nrow(xy))
```

```{r}

```

#want to CV errors for each step of forward selection. 
```{r,echo=FALSE}
# First we are going to load the cross-validation function
cvFolds = function (n, K = 5, R = 1, type = c("random", "consecutive", "interleaved")) 
{
  n <- round(rep(n, length.out = 1))
  if (!isTRUE(n > 0)) 
    stop("'n' must be positive")

  K <- round(rep(K, length.out = 1))
  if (!isTRUE((K > 1) && K <= n)) 
    stop("'K' outside allowable range")

  type <- if (K == n) 
    "leave-one-out"
  else match.arg(type)
  if (type == "random") {
    R <- round(rep(R, length.out = 1))
    if (!isTRUE(R > 0)) 
      R <- 1
    subsets <- replicate(R, sample(n))
  }
  else {
    R <- 1
    subsets <- as.matrix(seq_len(n))
  }
  which <- rep(seq_len(K), length.out = n)
  if (type == "consecutive") 
    which <- rep.int(seq_len(K), tabulate(which))
  folds <- list(n = n, K = K, R = R, subsets = subsets, which = which)
  class(folds) <- "cvFolds"
  folds
}
```

Calculate the CV error for glms using model selection via backward selection.
```{r,echo=FALSE}
# The following function does cross-validation using
# X - predictor matrix
# y - class matrix
# V - number of CV folds
# seed - (optional) internally sets the seed.
cv.glm.backward = function(X,y,V,seed=NA,pen)
{
  # Set the seed
  if (!is.na(seed)) {
    set.seed(seed)
  }
  
  # Set n
  n = length(y)
  
  # Split the data up into V folds
  cvSets <- cvFolds(n, V)
  
  # Loop through each fold and calculate the error for that fold
  test.error <- c()
  for (i in 1:V) 
  {
    # set the indices corresponding to the training and test sets
    testInds <- cvSets$subsets[which(cvSets$which==i)]
    trainInds <- (1:n)[-testInds]
    
    X = data.frame(X)
    
    # Separate y and X into the training and test sets
    y.test <- y[testInds]
    X.test <- X[ testInds,]
    y.train <- y[trainInds]
    X.train <- X[trainInds,]
    
    #xy<-na.omit(cbind(X.train,y.train))
    #y.train0<-xy[,ncol(xy)]
    #X.train0<-xy[,-ncol(xy)]
    
    # Do classification on ith fold
    full <- lm(y.train ~., data=X.train)
    res.step <- step(full,k=pen)
    res = round(predict(res.step, newdata=X.test, type="response"))
    
    # Calcuate the test error for this fold
    test.error[i] <- sum(res!=y.test)
  }
  
  # Calculate the mean error over each fold
  cv.error = sum(test.error)/n
  
  # Return the results
  return(cv.error)
}
```

Calculate the CV error for glms using AIC and BIC
```{r,echo=FALSE,results='hide'}
data1 = na.omit(data)
y2 = data1$BMISC
X2 = data1[c("BDYMSQ04","BMR","SEX","HYPBC","EIBMR1","ADTOTSE")]
V=10
N1 = nrow(X2)
res.glm.cv.back.BIC = cv.glm.backward(X2,y2,V,seed=1,pen=log(N1))
res.glm.cv.back.AIC = cv.glm.backward(X2,y2,V,seed=1,pen=2)
```

###ANOVA
```{r}
m1_aov = aov(BMISC ~ 1 + BDYMSQ04 + SEX, data = proj_dat_diet2)
par(mfrow=c(2,2)) 
plot(m1_aov) 
#assumptions NOT met for ANOVA between BMI wrt diet and sex except for normality in residuals (QQ plot)
```

###Simple linear regression
```{r}
hist(log(proj_dat_diet2$BMISC))
plot(BMISC~BMR, data = proj_dat_diet2)
lm.out = lm(log(BMISC) ~ BMR, data = proj_dat_diet2) #continue to use sqrt transformed data throughout?
summary(lm.out)

#checkl assumptions for untransformed data - assumtpions not met
par(mfrow=c(2,2)) 
boxplot(lm.out$residuals) 
plot(lm.out, which=1:3, add.smooth=FALSE) 

#assumptions for log trasnformed data- assumptiosn appear better met than for non- transformed
plot(log(BMISC)~log(BMR), data = proj_dat_diet2)
lm.out1 = lm(log(BMISC) ~ log(BMR), data = proj_dat_diet2) 
par(mfrow=c(2,2)) 
boxplot(lm.out1$residuals) 
plot(lm.out1, which=1:3, add.smooth=FALSE)

par(mfrow=c(2,2)) 
plot(log(proj_dat_diet2$BMISC),log(proj_dat_diet2$BMR), pch=19) 
abline(coef(lm.out1)) 
plot(lm.out1, which=c(1,2,3), add.smooth=TRUE)

```

Multiple linear regression
```{r}
lapply(proj_dat_diet2, class)
lm.mult = lm(BMISC ~ AGEC + SEX + BDYMSQ04, data = proj_dat_diet2)
summary(lm.mult)

#check assumptions - are not met
par(mfrow=c(2,2)) 
boxplot(lm.mult$residuals) 
plot(lm.mult, which=1:3, add.smooth=FALSE)
```



