---
output:
  pdf_document: default
  html_document: default
---
```{R} 
setwd("C:/Users/JPRS1/Desktop/STAT3014/Major project/STAT3014-Major-Project")
proj_dat = read.csv("cleanedData.csv",row.names=1)
dataSteph = proj_dat[which(proj_dat$AGEC >= 18), ]

carb = cut(dataSteph$CHOPER1, breaks=c(-1, 45, 65, 100), labels=c("low", "med", "high"))
protein = cut(dataSteph$PROPER1, breaks=c(-1, 15, 25, 100), labels=c("low", "med", "high"))
fat = cut(dataSteph$FATPER1, breaks=c(-1, 20, 35, 100), labels=c("low", "med", "high"))

library(MASS)
#bmi, age, exercise, 
dat1.var = dataSteph[,c(1,2,8, 11, 12, 14, 77, 86, 64, 65, 68)]
names(dat1.var) = c("bmi", "age", "mins.phys", "waist.cm", "bmr", "ses", "mins.sed", "sex", "protein", "fat", "carbs")
```

#can we predict who is obese according to BMI?
```{R}
bmi.class = cut(dat1.var$bmi, breaks=c(18.5, 30, 65), labels=c("norm","obese")) #categorical
bmi.num = ifelse(bmi.class=="norm", 0,1) #binary
bmi1 = dat1.var$bmi #numeric
# lot's of NA's
X = data.matrix(dat1.var[2:8])

#obese as binary for variables
mod = glm(bmi.num~X, family = binomial)
summary(mod)

#obese as continuous for variables
mod1 = glm(bmi1~X)
summary(mod1)

```
Have a smaller AIC when the data is run as continuous instead of binary. Using the binary variable model we have a log linear model of
$log(p/1-p) = -43.3 + 0.05*age + 0.18*waist.cm + 0.002*bmr - 0.00*mins.sed + 5.26*sex$
This tells us that the best model for describing BMI includes the following relationships:
- Higher age have a higher BMI
- Larger waist measurements have higher BMI
- Higher BMR have higher BMI
- More minutes per week sedentary is related to a higher BMI
- Females have a higher BMI than males

#Part 3: want to see if people of different diets have different behaviours and demographics. Use deviance tests 

$H_0: null model (intercept only)$
$H_1: one factor model, protein$

In both hypothesis testing using the AIC and the deviance test we failed to reject the null hypothesis and prefer the one factor model for protein. BMI is only significant with respect to protein. To fit the mean percentage of obesity.  
Part 3 - Model Selection using AIC
Stepwise forward regression was used to select the most informative variables, which were included in a generalised linear model (GLM). GLMs were used because they are able to handle different types of data including binary, categorical, and numerical. Logistic regression was used for binary data and classical regression was used for continuous data. A 5% significance level was chosen as a threshold for the inclusion of the model variables.
In this section the aim was to determine the effectiveness of the three macro- nutrients for predicting gender, socio-economic status, Basal Betabolic Rate (BRM), BMI, age, and mintues spent sedentary for subjects in this study. These variables were chosen because they were found to be significant predictors of BMI. Can we determine the type of diet someone may have according to their SES, age, gender, waist measurement, bmr, time spent sedentary, energy intake, and time spend doing exercise. 



BMI - BMISC
```{r}
bmi.class = cut(dat1.var$bmi, breaks=c(18.5, 30, 65), labels=c("norm","obese")) #categorical
bmi.num = ifelse(bmi.class=="norm", 0,1) #binary

bmi.macro = glm(bmi.num~protein+carb+fat, family = binomial)
best.marco = stepAIC(bmi.macro, scope = list(upper = ~protein*carb*fat, lower = ~1))
summary(best.marco)

#deviance test
mod12 = glm(bmi.num~protein, family = binomial)
mod13 = glm(bmi.num~fat, family = binomial)
mod14 = glm(bmi.num~carb, family = binomial)
a = anova(mod12, test = "Chisq")
b = anova(mod13, test = "Chisq")
c = anova(mod14, test = "Chisq")
a #protein is significant. So now look for protein and fat and protein and carb. 
b
c

mod15 = glm(bmi.num~protein+fat, family = binomial) #not signif
mod16 = glm(bmi.num~protein+carb, family = binomial) #not signif
d = anova(mod15, test = "Chisq")
e = anova(mod16, test = "Chisq")
d
e

pars = coef(best.marco)
logistic = function(x){1/(1+exp(-x))}
prop.low = logistic(pars[1])
prop.med = logistic(pars[1]+pars[2])
prop.high = logistic(pars[1]+pars[2]+pars[3])
```
AIC: bmi.num ~ protein
Deviance:  bmi.num ~ protein
The estimated proprotion of obese people within the low protein group is 24.6% (since only working with intercept). 
The estimated proprotion of obese people within the medium protein group is 28.2%. 
The estimated proprotion of obese people within the high protein group is 40.1%.

Waist measurement
```{r}
#AIC model selection
waist = dat1.var$waist.cm
aovwaist = lm(waist~protein+fat+carb)
bestmod.waist = stepAIC(aovwaist , scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.waist)

#deviance test
mod17 = lm(waist~protein)
mod18 = lm(waist~fat)
mod19 = lm(waist~carb)
a = anova(mod17)
b = anova(mod18)
c = anova(mod19)
a #protein is significant and explains the most of the within factor variation out of the three variables. 
b
c 

#protein
mod20 = lm(waist~protein+fat) 
mod21 = lm(waist~protein+carb) #signif
d = anova(mod20)
e = anova(mod21)
d
e


mod21 = lm(waist~protein+carb+fat) #not signif
f = anova(mod21)
f

#mean waist size 
most.ext = 93.3089
mn.high.pro = 93.3089 +1.8730
mn.med.carb = 93.3089 +1.8730 - 0.895

#using predict()

```
AIC: waist ~ protein + carb
Deviance: waist~protein+carb
Because were are now doing normal regression we are looking at the mean within each group, not the proportion. The largest mean waist size in this model was found to be 95.12cm for people with a high protein diet. 


BMR
```{r}
bmr = dat1.var$bmr
aovbmr = lm(bmr~protein+fat+carb)
bestmod.bmr = stepAIC(lm(bmr~protein*fat*carb) , scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.bmr)

#deviance test
mod17 = lm(bmr~protein) #significant at 0.1 level
mod18 = lm(bmr~fat)
mod19 = lm(bmr~carb)
a = anova(mod17)
b = anova(mod18)
c = anova(mod19)
a  
b
c


```
AIC: bmr ~ protein + fat + carb + protein:fat + protein:carb + fat:carb
Deviance: bmr~protein
Choose AIC since the AIC has a lower F statistic than the Devicance test. 

Sedentary
```{r}
sed = dat1.var$mins.sed
aovsed = lm(sed~protein+fat+carb)
bestmod.sed = stepAIC(aovsed , scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.sed)

#deviance test
mod17 = lm(sed~protein)
mod18 = lm(sed~fat)
mod19 = lm(sed~carb)
a = anova(mod17)
b = anova(mod18)
c = anova(mod19)
a  
b #most signif
c

#fat
mod20 = lm(sed~fat+protein) #signif
mod21 = lm(sed~fat+carb) 
d = anova(mod20)
e = anova(mod21)
d
e

#fat+protein
mod20 = lm(sed~fat+protein+carb)
d = anova(mod20)
d #signif

#interaction
mod20 = lm(sed~fat*protein*carb) #not signif
d = anova(mod20)
d 

#mean of each group

```
AIC: sed ~ protein + fat + carb
Deviance: sed ~ protein + fat + carb


Sex
```{R}
#AIC
sex = ifelse(dat1.var$sex==1,1,0) #1 = male
glm.sex = glm(sex~protein+fat+carb, family= binomial)
bestmod.sex = stepAIC(glm.sex, scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.sex)

#deviance test
mod17 = lm(sex~protein, family = "binomial")
mod18 = lm(sex~fat,  family = "binomial")
mod19 = lm(sex~carb,  family = "binomial")
a = anova(mod17, test = "Chisq")
b = anova(mod18, test = "Chisq")
c = anova(mod19, test = "Chisq")
a 
b #signif: explains more within SS variation
c  #signif

#fat
mod20 = glm(sex~fat+protein, family = "binomial") #not signif
mod21 = glm(sex~fat+carb, family = "binomial") #signif
d = anova(mod20, test = "Chisq")
e = anova(mod21, test = "Chisq")
d
e

#fat+carb
mod22 = glm(sex~fat+carb+protein, family = "binomial") #signif
f = anova(mod22, test = "Chisq") 
f

fat[which.max(glm.sex$fitted.values)]
protein[which.max(glm.sex$fitted.values)]
carb[which.max(glm.sex$fitted.values)]
max(glm.sex$fitted.values)

```
AIC: sex ~ protein + fat + carb + protein:fat
Deviance: sex ~ protein + fat + carb + protein:fat
The expected proportion of males in the the low fat, low protein, low carb group is 59%. 

SES
```{r}
dataSteph$SF2SA1QN = as.numeric(as.factor(dataSteph$SF2SA1QN))
ses = ifelse(dataSteph$SF2SA1QN>4,1,0) #decile [1:4] = 0
glmadd.ses = glm(ses~protein+fat+carb, family= binomial)
bestmod.ses = stepAIC(glmadd.ses, scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.ses)

mod23 = glm(ses~protein, family = "binomial")
mod24 = glm(ses~fat,  family = "binomial")
mod25 = glm(ses~carb,  family = "binomial") 
a = anova(mod23, test = "Chisq")
b = anova(mod24, test = "Chisq")
c = anova(mod25, test = "Chisq")
a 
b 
c #signif

#carbs
mod20 = glm(sex~carb+protein, family = "binomial") 
mod21 = glm(sex~carb+fat, family = "binomial") #most signif
d = anova(mod20, test = "Chisq")
e = anova(mod21, test = "Chisq")
d
e

#fat+carb
mod22 = glm(sex~carb+fat+protein, family = "binomial") #signif
f = anova(mod22, test = "Chisq") 
f

#carb*fat*protein
mod23 = glm(sex~carb*fat*protein, family = binomial)
g = anova(mod23, test = "Chisq")
g

#percentages
fat[which.max(glmadd.ses$fitted.values)]
protein[which.max(glmadd.ses$fitted.values)]
carb[which.max(glmadd.ses$fitted.values)]
max(glmadd.ses$fitted.values)
```
AIC: ses ~ protein + fat + carb + protein:fat
Deviance: ses ~ protein + fat + carb + protein:fat
In the low fat, medium protein, low carb diet a person from medium SES will be in this diet with a probability of 26%. 

Age - using ANOVA since age is not a factor but a numercial variable. 
```{r}
#anova - check if model good or not
#lm - estimate coefficients use summary(lm()) for model
age = dataSteph$AGEC
aovadd.age = lm(age~protein+fat+carb)
bestmod.age = stepAIC(aovadd.age, scope = list(upper = ~protein*fat*carb, lower = ~1))
summary(bestmod.age)

mod17 = lm(age~protein) #signif
mod18 = lm(age~fat)
mod19 = lm(age~carb)
a = anova(mod17)
b = anova(mod18)
c = anova(mod19)
a  
b 
c

#protein
mod18 = lm(age~protein+carb) #signif
mod19 = lm(age~protein+fat)
d = anova(mod18)
e = anova(mod19)
d
e

mod20 = lm(age~protein+carb+fat) #signif
f = anova(mod20)
f


mod21 = lm(age~protein*carb*fat) #not signif
g = anova(mod21)
g

#expected mean
fat[which.max(aovadd.age$fitted.values)]
protein[which.max(aovadd.age$fitted.values)]
carb[which.max(aovadd.age$fitted.values)]
max(aovadd.age$fitted.values)
```
AIC: age ~ protein + fat + carb
Deviance: age ~ protein + fat + carb
The most extreme value is in the low fat, medium protein, low carb group. The mean age in this group is 52 years. 









